---
title: "Video preparation and tracking with `idtrackerai` on Google Cloud"
date: '`r format(Sys.Date())`'
output:
  bookdown::tufte_html_book:
    css: "my_style.css"
    config:
      toc:
    pandoc_args: --lua-filter=color-text.lua
    highlight: pygments
    link-citations: yes
---

# Video preparation and tracking

Link to `Medaka Cloud Pilot` project on Google Cloud: <https://console.cloud.google.com/home/dashboard?project=medaka-cloud-pilot>.

Storage here in `medaka-video-store`: <https://console.cloud.google.com/storage/browser/medaka-video-store;tab=objects?forceOnBucketsSortingFiltering=false&project=medaka-cloud-pilot&prefix=&forceOnObjectsSortingFiltering=false>.

## Get R source

```{r, warning = F, message=F}
source(here::here("docs/01_tracking_source.R"))
```


## Video processing

Carried out with Snakemake:



## Full script to process in GCP terminal

```{bash, eval = F}
# Create `videos` directory
mkdir videos

# Clone repo
git clone https://github.com/brettellebi/pilot_paper.git

# Install `gsutil`
curl -O https://dl.google.com/dl/cloudsdk/channels/rapid/downloads/google-cloud-sdk-327.0.0-linux-x86_64.tar.gz
tar -xvf google-cloud-sdk-327.0.0-linux-x86_64.tar.gz
./google-cloud-sdk/install.sh
rm google-cloud-sdk-327.0.0-linux-x86_64.tar.gz

## Create authorisation
gcloud auth login

# Install `conda` 
wget https://repo.anaconda.com/miniconda/Miniconda3-latest-Linux-x86_64.sh
chmod +x Miniconda3-latest-Linux-x86_64.sh
./Miniconda3-latest-Linux-x86_64.sh
rm Miniconda3-latest-Linux-x86_64.sh

# Install `idtrackerai`
conda create -y -n idtrackerai python=3.6
conda activate idtrackerai
pip install idtrackerai[gpu]

## Copy processed results folder for training
### icab
mkdir -p trained/icab
gsutil -m cp -r $video_store/results/session_20190615_1649_icab_icab_L_q2_open_field/accumulation_0/* trained/icab
### hdr
mkdir -p trained/hdr
gsutil -m cp -r $video_store/results/session_20190615_1452_icab_hdr_L_q2_novel_object/accumulation_0/* trained/hdr
### hni
mkdir -p trained/hni
gsutil -m cp -r $video_store/results/session_20190616_1325_icab_hni_R_q2_novel_object/accumulation_0/* trained/hni
### kaga
mkdir -p trained/kaga
gsutil -m cp -r $video_store/results/session_20190616_1017_icab_kaga_R_q1_novel_object/accumulation_0/* trained/kaga
### ho5
mkdir -p trained/ho5
gsutil -m cp -r $video_store/results/session_20190616_1017_icab_kaga_R_q1_novel_object/accumulation_0/* trained/ho5


# Script
## Copy processed results folder for training
gsutil -m cp -r gs://medaka-video-store/results/session_20190615_1649_icab_icab_L_q2_open_field/accumulation_0 .

## Select number of random samples to process successively
samples_no=5
## Sample parameters file
samples_file=pilot_paper/data/20210118_paramaters.csv
## Video store
video_store=gs://medaka-video-store

## Loop script
# Alternative first lines:
for in_sample in $(gsutil ls $video_store/videos | cut -f5 -d'/' | cut -f1 -d'.' | shuf -n $samples_no); do

for in_sample in $(gsutil ls $video_store/videos/2019061{1,2,3}* | cut -f5 -d'/' | cut -f1 -d'.' ); do

for in_sample in $(gsutil ls $video_store/videos/20190614* | cut -f5 -d'/' | cut -f1 -d'.' ); do

for in_sample in $(gsutil ls $video_store/videos/2019061{5,6}* | cut -f5 -d'/' | cut -f1 -d'.' ); do
for in_sample in $(gsutil ls $video_store/videos/2019061{1,2,3}* | cut -f5 -d'/' | cut -f1 -d'.' ); do
for in_sample in $(gsutil ls $video_store/videos/20190614* | cut -f5 -d'/' | cut -f1 -d'.' ); do
  # check whether final trajectories_wo_gaps.npy file exists in results folder
  gsutil -q stat $video_store/results/session_$in_sample/trajectories_wo_gaps/trajectories_wo_gaps.npy ;
  status=$? ;
  # Run `idtrackerai` if results don't exist
  if [[ $status == 1 ]]; then
    # Create log
    echo "$in_sample: started $(date)" >> pilot_paper/data/20210211_test_train.txt ;
    # Set training folder
#    target_line=$(echo $in_sample | cut -f4 -d'_' )
#    echo "KNOWLEDGE_TRANSFER_FOLDER_IDCNN = trained/$target_line" > local_settings.py ;
    echo "KNOWLEDGE_TRANSFER_FOLDER_IDCNN = accumulation_0" > local_settings.py ;
    # Copy video to VM
    gsutil -m cp $video_store/videos/$in_sample.mp4 videos/ ;
    # Get parameters
    input_video=$(echo videos/$in_sample.mp4 ) ;
    target_string=$( grep $in_sample $samples_file ) ;
    vid_length=$( echo $target_string | cut -f2 -d',' ) ;
    int_floor=$( echo $target_string | cut -f3 -d',' ) ;
    int_ceiling=$( echo $target_string | cut -f4 -d',' ) ;
    area_floor=$( echo $target_string | cut -f5 -d',' ) ;
    area_ceiling=$( echo $target_string | cut -f6 -d',' ) ;
    # Run `idtrackerai`
    idtrackerai terminal_mode \
              --_video $input_video \
              --_bgsub 'True' \
              --_intensity [$int_floor,$int_ceiling] \
              --_area [$area_floor,$area_ceiling] \
              --_range [0,$vid_length] \
              --_nblobs 2 \
              --_session $in_sample \
              --exec track_video ;
    # Run script to convert tracking data to CSVs
    python pilot_paper/code/scripts/trajectories_to_csv.py videos/session_$in_sample ;
    # Copy to storage bucket
    gsutil -m cp -r \
      videos/session_$in_sample \
      gs://medaka-video-store/results ;
    # Remove video and results
    rm videos/$in_sample.mp4
    rm -rf videos/session_$in_sample
    # Confirm with output
    echo "$in_sample: completed $(date)" >> pilot_paper/data/20210211_test_train.txt
    echo "$in_sample: VIDEO PROCESSED AND COPIED TO BUCKET. PROCEEDING TO NEXT..."
  else
    echo "$in_sample: VIDEO ALREADY PROCESSED. PROCEEDING TO NEXT..."
  fi
done
```

## Copy back over to Codon HPC

```{bash, eval = F}
gsutil -m cp -r gs://medaka-video-store/results/* /nfs/research/birney/users/ian/pilot/tracking_results/
```

## Process tracking data

### Read in data

#### Samples files

```{r}
samples_file = here::here("code/snakemake/20210701/config/samples.csv")
samples_pairs_file = here::here("code/snakemake/20210701/config/samples_pairs.csv")

samples = readr::read_csv(samples_file,
                          col_types = c("ciiciiiiiiiiiiiiiiicccccccc"))

samples_pairs = readr::read_csv(samples_pairs_file,
                                col_types = c("ciiiiic"))
```

#### Add video height and width

```{r}

```


#### Tracking data

```{r}
in_dir = here::here("data/tracking_round_01")

target_files = list.files(in_dir, full.names = T)

dat_list = purrr::map(target_files, function(FILE){
  out = list()
  # Read in tracking data
  out[["RAW"]] = readr::read_csv(FILE,
                        col_types = c("dddd"))
  colnames(out[["RAW"]]) = c("X1", "Y1", "X2", "Y2")

  # Add metadata
  ## Parse file name
  parsed_meta = basename(FILE) %>% 
    stringr::str_remove(".csv") %>% 
    stringr::str_split("_", simplify = T)
  ## Add to data frame
  out[["META"]] = tibble::tibble(VIDEO = paste(parsed_meta[, 1:5], collapse = "_"),
                                 DATE = as.integer(parsed_meta[, 1]),
                                 TIME = as.integer(parsed_meta[, 2]),
                                 REF = parsed_meta[, 3],
                                 TEST = parsed_meta[, 4],
                                 TANK_SIDE = parsed_meta[, 5],
                                 QUADRANT = parsed_meta[, 6],
                                 ASSAY = paste(parsed_meta[, 7], parsed_meta[, 8], sep = "_"))
  ## Add frame rate
  out[["META"]] = dplyr::left_join(out[["META"]],
                                   samples %>% 
                                     dplyr::select(date, time, fps),
                                   by = c("DATE" = "date", "TIME" = "time")) %>% 
    dplyr::rename(FPS = fps)
  ## Add iCab location
  out[["META"]]$ICAB_LOC = samples_pairs %>% 
    dplyr::filter(sample == basename(FILE) %>% 
                    stringr::str_remove(".csv")) %>% 
    dplyr::pull(icab_loc)
  
  # Create new data frame with only frames up to 10 minutes
  frame_cutoff = out[["META"]]$FPS * 60 * 10
  out[["TRACKING"]] = out[["RAW"]][1:frame_cutoff, ]
  
  return(out)
})
names(dat_list) = basename(target_files) %>% 
        stringr::str_remove(".csv")

```

## How many NAs?

```{r}
na_df = purrr::map(dat_list, function(SAMPLE){
  out = tibble::tibble(N = nrow(SAMPLE$TRACKING),
                       N_NA = SAMPLE$TRACKING %>% 
                         dplyr::filter(!complete.cases(.)) %>% 
                         nrow(.)) %>% 
    dplyr::mutate(PROP_NA = N_NA / N )
  return(out)
}) %>% 
  dplyr::bind_rows(.id = "SAMPLE")



# Plot

na_plot = na_df %>% 
  ggplot() +
    geom_histogram(aes(N_NA), bins = 200)

ggplotly(na_plot)
```

### Filter out from data list the videos with more than 10% NAs

```{r}
to_filter_out = na_df %>% 
  dplyr::filter(N_NA > 1800) %>% 
  dplyr::pull(SAMPLE)

dat_filt = dat_list[!(names(dat_list) %in% to_filter_out)]

# Get rid of raw data
dat_filt = dat_filt %>% 
  purrr::map(., function(SAMPLE) {
    purrr::keep(SAMPLE, names(SAMPLE) %in% c("META", "TRACKING"))
  })

```

### Plot tracking data

With assistance from here: https://github.com/ropensci/plotly/issues/957

```{r}
assays = c("open_field", "novel_object")
# Loop over each video
test3 = purrr::map(samples$sample[1:2], function(VIDEO){
  # Loop over each assay
  assays_out = purrr::map(assays, function(ASSAY){
    # Select the videos from each video and assay
    target_vids = purrr::keep(dat_filt, function(PAIR){
      PAIR[["META"]]$VIDEO == VIDEO & PAIR[["META"]]$ASSAY == ASSAY
    })
  })
  names(assays_out) = assays
  return(assays_out)
})

quadrants = c("q1", "q2", "q3", "q4")

test4 = test3[[1]][[2]]

plot_list = purrr::map(quadrants, function(QUADRANT){
  
})


test = dat_filt$`20190611_1331_icab_icab_R_q1_novel_object`
FPS = dat_filt$`20190611_1331_icab_icab_R_q1_novel_object`$META$FPS

recode_vec = label_fishes(test)

test2 = list(A = test$TRACKING %>% 
               dplyr::select(COORD_X = X1,
                             COORD_Y = Y1) %>% 
               dplyr::mutate(FRAME = seq(1:nrow(.))),
             B = test$TRACKING %>% 
               dplyr::select(COORD_X = X2,
                             COORD_Y = Y2) %>% 
               dplyr::mutate(FRAME = seq(1:nrow(.)))) %>% 
  dplyr::bind_rows(.id = "FISH") %>% 
  # add line for palette
  dplyr::mutate(LINE = dplyr::recode(FISH, !!!recode_vec)) %>% 
  # order for plotting
  dplyr::mutate(LINE = factor(LINE, levels = names(fish_pal))) %>% 
  # add `SECONDS` and `TIME` columns
  dplyr::mutate(SECONDS = FRAME / FPS,
                TIME = hms::hms(SECONDS) %>%
                  strftime("%M:%S")) %>% 
  # keep first frame from every second for each fish
  dplyr::distinct(TIME, FISH, .keep_all = T) %>% 
  dplyr::filter(FRAME < 1000)

tracking_plot = function(input){
  # Create long DF 
  df = input %>%
    accumulate_by(~TIME)
  # Generate plot
  out_plot = df %>% 
    plot_ly(x = ~COORD_X, y = ~COORD_Y) %>% 
    add_paths(color = ~LINE,
              colors = fish_pal,
              line = list(simplify = F),
              frame = ~time) %>% 
    layout(yaxis = list(autorange = "reversed")) %>% 
    layout(xaxis = list(scaleanchor = "y", scaleratio = 1))
  
  return(out_plot)
}

tracking_plot(test2)

df = test2 %>%
  accumulate_by(~TIME)

df %>% 
  plot_ly(x = ~COORD_X, y = ~COORD_Y) %>% 
  add_paths(color = ~LINE,
            colors = fish_pal,
            line = list(simplify = F),
            frame = ~time) %>% 
  layout(yaxis = list(autorange = "reversed")) %>% 
  layout(xaxis = list(scaleanchor = "y", scaleratio = 1))
```

## Try with `gganimate`

```{r}
test = dat_filt$`20190611_1331_icab_icab_R_q1_novel_object`$TRACKING

test2 = list(A = test %>% 
               dplyr::select(COORD_X = X1,
                             COORD_Y = Y1) %>% 
               dplyr::mutate(FRAME = seq(1:nrow(.))),
             B = test %>% 
               dplyr::select(COORD_X = X2,
                             COORD_Y = Y2) %>% 
               dplyr::mutate(FRAME = seq(1:nrow(.)))) %>% 
  dplyr::bind_rows(.id = "FISH") %>% 
  # select every 30th frame
  dplyr::filter(FRAME %in% seq(from = 1, to = nrow(test), by = 30))

test2 %>% 
#  dplyr::filter(FRAME < 100) %>% 
  ggplot() +
    geom_path(aes(COORD_X, COORD_Y, colour = FISH)) +
    xlim(0, max(test2$COORD_X)) +
    scale_y_reverse(limits = c(max(test2$COORD_Y), 0)) +
    coord_fixed() +
    gganimate::transition_reveal(FRAME, keep_last = F)
```

