# HMM parameter optimisation {#param-optim}

## Setup

### Load libraries

```{r, warning=F, message = F}
library(tidyverse)
library(cowplot)
```

### Set variables

```{r}
CONC = as.list(list.files("/hps/nobackup/birney/users/ian/pilot/hmm_concordance_recode/",full.names = T, recursive = T))
KW = "/hps/nobackup/birney/users/ian/pilot/kruskal_wallis/out.rds"
OUT_PNG = here::here("book/figs/compare_params/compare_params.png")
OUT_PDF = here::here("book/figs/compare_params/compare_params.pdf")
```

## Read data and process

```{r}
kw = readRDS(KW) %>% 
  dplyr::mutate(INTERVAL = as.numeric(INTERVAL)) %>% 
  dplyr::mutate(N_STATES = as.numeric(N_STATES))

conc = purrr::map_dfr(CONC, readRDS) %>% 
  dplyr::group_by(INTERVAL, N_STATES) %>% 
  # get mean across groups A and B
  dplyr::summarise(MEAN_CONC = mean(CONCORDANCE))

df = dplyr::left_join(kw, conc, by = c("INTERVAL", "N_STATES"))
```

## Plot

```{r}
pal = colorspace::sequential_hcl(length(unique(df$N_STATES)), palette = "ag_Sunset")

out_plot = df %>% 
  dplyr::mutate(N_STATES = factor(N_STATES, levels = sort(unique(N_STATES)))) %>% 
  dplyr::mutate(INTERVAL = factor(INTERVAL, levels = sort(unique(INTERVAL)))) %>% 
  # make new column combining `INTERVAL` and `N_STATES`
  tidyr::unite(INTERVAL, N_STATES,
               col = "INT_STATE",
               sep = ";",
               remove = F) %>% 
  ggplot(aes(MEAN_CONC, SUM_KW_STAT)) +
    geom_point(aes(size = INTERVAL, colour = N_STATES),
               alpha = 0.8) +
    ggrepel::geom_text_repel(aes(label = INT_STATE),
              size = 2,
              ) +
    theme_bw() +
    guides(size = guide_legend(title = "Interval\n(seconds)"),
           colour = guide_legend(title = "N states")) +
    scale_colour_manual(values = pal) +
    xlab("mean concordance between cross-validated HMM states") +
    ylab("summed Kruskal-Wallis statistic comparing frequency\nof time spent in each HMM state across medaka lines")

plotly::ggplotly(out_plot)
```

### Save

```{r, eval = F}
ggsave(OUT_PNG,
       out_plot,
       device = "png",
       width = 10,
       height = 8.5,
       units = "in",
       dpi = 400)

ggsave(OUT_PDF,
       out_plot,
       device = "pdf",
       width = 10,
       height = 8.5,
       units = "in",
       dpi = 400)
```

### Show paper figure

(ref:hmm-param) Comparison between HMM parameters. Horizontal axis: Mean concordance between states assigned by HMMs through a 2-fold cross-validation process. Vertical axis: Kruskal-Wallis statistic comparing lines based on the proportion of time spent in each HMM state, summed across all states. Size of points correspond to the interval, in seconds, between which the distance and angle of travel was calculated.

```{r hmm-param, fig.cap = '(ref:hmm-param)', out.width='100%'}
knitr::include_graphics(OUT_PNG)
```

