# Time dependence of HMM states (DGE)

**Figure \@ref(fig:time-dep-dge)** depicts the time dependence of HMM states over the course of the video, and the regions of the test tank that were most frequently occupied by the different lines. There is an especially clear difference between lines in the proportions of time spent in the slowest-moving states (i.e. states 1 to 3) at the beginning of each assay component, with an increase across all lines in the novel object assay, likely as a consequence of having less room to move, as well as the fear response that the novel object was designed to elicit. The differences are clearest when comparing the southern Japanese medaka lines (*iCab*, *HdrR*, and *HO5*) against the northern Japanese lines (*Kaga* and *HNI*). 

In addition, we note that *Kaga* tends to spend more time in the fast- and forward-moving state 13 at the beginning of the assay than at other times, which suggests that for that line, such movements may be the manifestation of a stress response. Figures **\@ref(fig:time-dep-dge)D** and **\@ref(fig:time-dep-dge)G** show, as densities, the regions of the tank that were most frequently occupied by each line. Although the northern Japanese lines *Kaga* and *HNI* were similarly fast-moving relative to the southern Japanese lines, in the open field assay component they appear to favour different regions of the tank – where *HNI* occupied the central regions of the tank with more frequency, *Kaga* tended to prefer swimming along the boundaries of the tank.

## Setup

### Load libraries

```{r, message = F, warning = F}
library(tidyverse)
library(cowplot)
library(ggridges)
library(viridisLite)
library(googlesheets4)
```

### Set variables

```{r}
IN = "/hps/nobackup/birney/users/ian/pilot/hmm_out/0.08/dist_angle/14.csv"
AOV_SHEET = "https://docs.google.com/spreadsheets/d/1_l72BZkmWyNAOfCUI8WGP4UfQuIPQtPZZmlRjQffvEs"
N_STATES = 15
POLAR_ALL_DGE_SIG_OF = here::here("book/figs/paper_final/0.08/dist_angle/14/polar_all_dge_sig_of.png")
POLAR_ALL_DGE_SIG_NO = here::here("book/figs/paper_final/0.08/dist_angle/14/polar_all_dge_sig_no.png")
OUT_DGE = here::here("book/figs/time_dependence/dist_angle/0.08_14_dge.png")

# Create line recode vector
line_vec = c("iCab", "HdrR", "HNI", "Kaga", "HO5")
names(line_vec) = c("icab", "hdr", "hni", "kaga", "ho5")

```

## Read and process data

```{r}
df = readr::read_csv(IN) %>% 
  # recode angle to sit between 0 and 360
  dplyr::mutate(angle_recode = ifelse(angle < 0,
                                      180 + (180 + angle),
                                      angle))

# Recode states by mean distance

rank_df = df %>% 
  dplyr::group_by(state) %>% 
  dplyr::summarise(mean_dist = mean(distance)) %>% 
  # rank
  dplyr::arrange(mean_dist) %>% 
  dplyr::mutate(rank = 1:nrow(.))

recode_vec = rank_df %>% 
  dplyr::pull(rank)
names(recode_vec) = rank_df %>% 
  dplyr::pull(state)

# Recode `state`

df = df %>% 
  dplyr::mutate(state_recode = dplyr::recode(state, !!!recode_vec))

# Recode `assay`

df = df %>% 
  dplyr::mutate(assay = stringr::str_replace(assay, "_", " "),
                assay = factor(assay, levels = c("open field", "novel object")))

# Add `line`

df = df %>% 
  dplyr::mutate(line = dplyr::case_when(fish == "ref" ~ ref_fish,
                                        fish == "test" ~ test_fish)) %>% 
  # recode
  dplyr::mutate(line = dplyr::recode(line, !!!line_vec)) %>% 
  # factorise to order
  dplyr::mutate(line = factor(line, levels = line_vec))

# Recode `test_fish` and order

df = df %>% 
  dplyr::mutate(test_fish = dplyr::recode(test_fish, !!!line_vec),
                test_fish = factor(test_fish, levels = line_vec))

```


## Pull significant states from spreadsheets

```{r}
SIGS_DGE_OF = googlesheets4::read_sheet(AOV_SHEET, sheet = "DGE_OF") %>% 
  dplyr::filter(`p-value FDR-adj` < 0.05 & `Variable` == "line") %>% 
  dplyr::pull(State) %>% 
  as.integer()

SIGS_DGE_NO = googlesheets4::read_sheet(AOV_SHEET, sheet = "DGE_NO") %>% 
  dplyr::filter(`p-value FDR-adj` < 0.05 & `Variable` == "line") %>% 
  dplyr::pull(State)%>% 
  as.integer()
```

## Tile plots

```{r}

SEC_INT = 2

dge_tile_df = df %>% 
  # remove iCab ref fishes (because DGE compares test fishes)
  dplyr::filter(!(line == "iCab" & fish == "ref")) %>% 
  #dplyr::slice_sample(n = 1e6) %>% 
  # add `indiv` column
  tidyr::unite(date, time, quadrant, fish,
               col = "indiv",
               sep = "_") %>%
  # get top state per 2 seconds
  dplyr::mutate(seconds_bin = floor(seconds / SEC_INT)) %>% 
  dplyr::group_by(assay, indiv, line, seconds_bin) %>% 
  dplyr::count(state_recode) %>% 
  dplyr::slice_max(order_by = n, n = 1) %>% 
  dplyr::ungroup() %>% 
  # reverse order by `indiv` so that the earliest videos are at the top
  dplyr::arrange(indiv) %>% 
  # convert `seconds_bin` back to seconds
  dplyr::mutate(seconds = seconds_bin * SEC_INT)

# Open field

dge_tile_of = dge_tile_df %>% 
  dplyr::filter(assay == "open field") %>% 
  ggplot() +
  geom_tile(aes(seconds, indiv, fill = state_recode)) + 
  facet_grid(rows = vars(line), cols = vars(assay), scales = "free") +
  scale_fill_viridis_c() +
  scale_y_discrete(limits = rev) +
  guides(fill = "none") +
  ylab("individual fish") +
  cowplot::theme_cowplot(font_size = 12) +
  theme(strip.background = element_blank(),
        strip.text = element_text(face = "bold"),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.line.y = element_blank()) 

# Novel object

dge_tile_no = dge_tile_df %>% 
  dplyr::filter(assay == "novel object") %>% 
  ggplot() +
  geom_tile(aes(seconds, indiv, fill = state_recode)) + 
  facet_grid(rows = vars(line), cols = vars(assay), scales = "free") +
  scale_fill_viridis_c() +
  scale_y_discrete(limits = rev) +
  guides(fill = "none") +
  ylab("individual fish") +
  cowplot::theme_cowplot(font_size = 12) +
  theme(strip.background = element_blank(),
        strip.text = element_text(face = "bold"),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.line.y = element_blank()) 

```

## Density plots

```{r}
# Take viridis colours for significant states and add grey
pal_dge_of = viridisLite::viridis(n = N_STATES)
pal_dge_of = c(pal_dge_of[SIGS_DGE_OF], "#9da2ab")
names(pal_dge_of) = c(as.character(SIGS_DGE_OF), "other")

pal_dge_no = viridisLite::viridis(n = N_STATES)
pal_dge_no = c(pal_dge_no[SIGS_DGE_NO], "#9da2ab")
names(pal_dge_no) = c(as.character(SIGS_DGE_NO), "other")

ASSAY = "open field"
time_dens_dge_of = df %>% 
  dplyr::filter(assay == ASSAY) %>% 
  # remove iCab when paired with a different test fish
  dplyr::filter(fish == "test") %>% 
  # filter for target assay
  #dplyr::filter(assay == ASSAY) %>% 
  # recode state 
  dplyr::mutate(state_plot_recode = dplyr::case_when(state_recode %in% SIGS_DGE_OF ~ as.character(state_recode),
                                                     TRUE ~ "other"),
                state_plot_recode = factor(state_plot_recode, levels = c(as.character(SIGS_DGE_OF), "other"))) %>% 
  ggplot() +
  geom_density(aes(seconds, after_stat(count), fill = state_plot_recode),
               position = "fill") +
  facet_grid(rows = vars(line),
             cols = vars(assay)) + 
  scale_fill_manual(values = pal_dge_of) +
  cowplot::theme_cowplot(font_size = 12) +
  theme(strip.background = element_blank(),
        strip.text = element_text(face = "bold")) +
  guides(fill = "none") +
  scale_x_continuous(breaks = c(0,200,400,600)) +
  scale_y_continuous(breaks = c(0,0.5,1))

ASSAY = "novel object"
time_dens_dge_no = df %>% 
  dplyr::filter(assay == ASSAY) %>% 
  # remove iCab when paired with a different test fish
  dplyr::filter(fish == "test") %>% 
  # filter for target assay
  #dplyr::filter(assay == ASSAY) %>% 
  # recode state 
  dplyr::mutate(state_plot_recode = dplyr::case_when(state_recode %in% SIGS_DGE_NO ~ as.character(state_recode),
                                                     TRUE ~ "other"),
                state_plot_recode = factor(state_plot_recode, levels = c(as.character(SIGS_DGE_NO), "other"))) %>% 
  ggplot() +
  geom_density(aes(seconds, after_stat(count), fill = state_plot_recode),
               position = "fill") +
  facet_grid(rows = vars(line),
             cols = vars(assay)) + 
  scale_fill_manual(values = pal_dge_no) +
  cowplot::theme_cowplot(font_size = 12) +
  theme(strip.background = element_blank(),
        strip.text = element_text(face = "bold")) +
  guides(fill = "none") +
  scale_x_continuous(breaks = c(0,200,400,600)) +
  scale_y_continuous(breaks = c(0,0.5,1))

```

## Spatial density plots

```{r}
# Create density function

get_density <- function(x, y, ...) {
  dens <- MASS::kde2d(x, y, ...)
  ix <- findInterval(x, dens$x)
  iy <- findInterval(y, dens$y)
  ii <- cbind(ix, iy)
  return(dens$z[ii])
}


# OF

sdens_dge_df = df %>% 
  #dplyr::slice_sample(n = 1e5) %>% 
  # filter out iCab references when paired with a different line
  dplyr::filter(fish == "test") %>% 
  # group by line
  dplyr::group_by(assay, line) %>% 
  # get densities
  dplyr::mutate(density = get_density(x, y, n = 30)) %>% 
  dplyr::ungroup()

  # Plot
sdens_dge_of = sdens_dge_df %>% 
  ## take only states 1:4
  dplyr::filter(assay == "open field") %>% 
  ggplot() +
  geom_point(aes(x, y, colour = density),
             alpha = 0.1, size = 0.2) +
  facet_grid(cols = vars(assay), rows = vars(line)) +
  colorspace::scale_color_continuous_sequential(palette = "Mako", rev = F) +
  #scale_colour_viridis_c(option = "rocket") +
  cowplot::theme_cowplot(font_size = 12) +
  theme(aspect.ratio = 1,
        strip.background = element_blank(),
        strip.text = element_text(face = "bold")) +
  scale_x_continuous(breaks = c(0,150,300)) +
  scale_y_continuous(breaks = c(0,150,300)) +
  xlab("x") +
  ylab("y")

# NO

sdens_dge_no = sdens_dge_df %>% 
  dplyr::filter(assay == "novel object") %>% 
  # Plot
  ggplot() +
  geom_point(aes(x, y, colour = density),
             alpha = 0.1, size = 0.2) +
  #coord_polar() +
  facet_grid(cols = vars(assay), rows = vars(line)) +
  colorspace::scale_color_continuous_sequential(palette = "Mako", rev = F) +
  #scale_colour_viridis_c(option = "rocket") +
  cowplot::theme_cowplot(font_size = 12) +
  theme(aspect.ratio = 1,
        strip.background = element_blank(),
        strip.text = element_text(face = "bold")) +
  scale_x_continuous(breaks = c(0,150,300)) +
  scale_y_continuous(breaks = c(0,150,300)) +
  xlab("x") +
  ylab("y")

```

## Combine for final figure

```{r, eval = F}
final_dge_raw = cowplot::plot_grid(dge_tile_of +
                                 theme(strip.background.y = element_blank(),
                                       strip.text.y = element_blank(),
                                       axis.title.y = element_text(vjust=-5)),
                               time_dens_dge_of + 
                                 theme(strip.background.y = element_blank(),
                                       strip.text.y = element_blank()),
                               sdens_dge_of,
                               dge_tile_no +
                                 theme(strip.background.y = element_blank(),
                                       strip.text.y = element_blank(),
                                       axis.title.y = element_text(vjust=-5)),
                               time_dens_dge_no +
                                 theme(strip.background.y = element_blank(),
                                       strip.text.y = element_blank()),
                               sdens_dge_no,
                               nrow = 2, ncol = 3,
                               rel_widths = c(1,1,0.6,1,1,0.6),
                               align = "hv",
                               labels = c('B', 'C', 'D', 'E', 'F', 'G'))

final_dge_with_polar = ggdraw() +
  cowplot::draw_image(POLAR_ALL_DGE_SIG_OF,
                      x = 0, y = 0.8,
                      width = 1, height = 0.2) +
  cowplot::draw_plot(final_dge_raw,
                     x = 0, y = 0.2,
                     width = 1, height = 0.6) +
  cowplot::draw_image(POLAR_ALL_DGE_SIG_NO,
                      x = 0, y = 0,
                      width = 1, height = 0.2) +
  cowplot::draw_plot_label(c('A', 'H'),
                           x = c(0,0), y = c(1, 0.2),
                           size = 14)

ggsave(OUT_DGE,
       final_dge_with_polar,
       device = "png",
       width = 11.5,
       height = 20,
       units = "in",
       dpi = 400)
```

(ref:time-dep-dge) Differences between test fish lines in the HMM states they occupied during the open field (top) and novel object (bottom) assay components. **A** and **H**: 14 HMM states with panels coloured red to indicate significant differences between test fish lines in the proportion of time spent in those states during the separate assay components. **B** and **E**: Transitions between HMM states across time for each individual test fish, grouped by line. Tiles are coloured by the state most frequently occupied by each fish within 2-second intervals. **C** and **F**: Densities within each line for the occupation of states that significantly differed between lines (colour), with other states consolidated (grey). **D** and **G**: Densities of the test tank locations occupied by each line, calculated within 900 grid points (30x30).

```{r time-dep-dge, out.width = '100%'}
knitr::include_graphics(OUT_DGE)
```

